---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
data <- read.csv(unz("./activity.zip", "activity.csv"))
summary(data)
```

```{r}
# Massage data into steps taken per day
dailyData <- aggregate(steps ~ date, data, sum, na.rm=T)
summary(dailyData)
```

```{r}
# Massage data into steps taken per 5-minute interval
intervalData <- aggregate(steps ~ interval, data, sum, na.rm=T)
intervalMeans <- aggregate(steps ~ interval, data, mean, na.rm=T)
summary(intervalData)
summary(intervalMeans)
```

## What is mean total number of steps taken per day?

```{r}
print(sprintf("Steps taken total: %d", sum(data$steps, na.rm=T)))

print(sprintf("Mean steps taken per 5-minute period: %f", mean(data$steps, na.rm=T)))

print(sprintf("Median steps taken per 5-minute period: %f", median(data$steps, na.rm=T)))

print(sprintf("Mean steps taken per day: %f", mean(dailyData$steps, na.rm=T)))

print(sprintf("Median steps taken per day: %f", median(dailyData$steps, na.rm=T)))

hist(dailyData$steps)
```

## What is the average daily activity pattern?


```{r}
# Plot time-series of total steps across 5-minute intervals, across all days
plot(intervalData$steps ~ intervalData$interval, type="l")
```

```{r}
# Which interval # is the highest overall?
head(sapply(intervalData, sort, decreasing=T), n=1)
```

## Imputing missing values

```{r}
# How many 5-minute intervals have missing data?
sum(is.na(data$steps))
# How many valid records?
sum(!is.na(data$steps))
# Fill in missing data with median from that interval
data2 <- data
data2[is.na(data2$steps),]$steps <- intervalMeans$steps

# Summary of imputed data
summary(data2)

# Calculate mean, median, daily histogram with imputed data
imputedDailyData <- aggregate(steps ~ date, data2, sum, na.rm=T)
summary(imputedDailyData)
print(sprintf("Mean steps taken per day: %f", mean(imputedDailyData$steps, na.rm=T)))
print(sprintf("Median steps taken per day: %f", median(imputedDailyData$steps, na.rm=T)))
hist(imputedDailyData$steps)
```

The mean remains unchanged but the median has increased slightly by 1 step per day.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
data2$dayOfWeek <- factor(weekdays(as.Date(data2$date)))
levels(data2$dayOfWeek) <- c("wd", "wd", "we", "we", "wd", "wd", "wd")
summary(data2)

# Aggregate steps over weekends, weekdays
weekdayVsWeekendSteps <- aggregate(steps ~ dayOfWeek + interval, data2, sum, na.rm=T)
head(weekdayVsWeekendSteps)
```

```{r}
library(lattice)
# Plot timeseries of weekend, weekday steps, across interval identifier
xyplot(steps ~ interval | dayOfWeek, data=weekdayVsWeekendSteps, type="l", layout=c(1,2))
```

On weekends, steps throughout the afternoon are relatively constant, while there isn't much evening or morning activity. Weekdays show the most activity at mid-day.